---
title: "Course Project -1"
author: "Chandana"
date: "6 June 2018"
output:
  md_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## To analyse personal movement data

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

```{r Load & Clean Data}
activity<-read.csv("activity.csv")
head(activity)
summary(activity$steps)
activity$date<-as.Date(activity$date,"%Y-%m-%d")
```

## Total steps every day

You can also embed plots, for example:

```{r total steps, echo=TRUE}
steps<-with(activity,tapply(steps,date,sum, na.rm=T))
hist(steps, xlab= "Total Steps", main= " Total Steps Taken Every Day", col= "steel blue")
mean0<-round(mean(steps))
median0<-median(steps)
```

The median of total steps taken in a day is `r median0`, mean is `r mean0`.

## Variation observed in different intervals
```{r variation intervals}
intervalsteps<-aggregate(steps~interval, activity, FUN = mean)
plot(intervalsteps$interval, intervalsteps$steps, type="l", xlab= "Intervals",ylab="Average Steps",main="Average steps per interval",col="orange")
max<-subset(intervalsteps, intervalsteps$steps==max(intervalsteps$steps))
maximum<-max$steps
maxinterval<- max$interval
```

The interval with maximum number of steps is `r maxinterval`.
```{r NAs}
nNA<-sum(is.na(activity$steps))
```

The number of NAs in the data set is `r nNA`.

### To impute NAs
```{r impute}
toclean<-merge(activity,intervalsteps, by= "interval")
toclean<-toclean[order(toclean$date),]
names(toclean)<-c("interval", "steps", "date","meanin" )
i<-0
while(i<nrow(toclean)){
  i<-i+1
  if(is.na(toclean$steps[i])){
    toclean$steps[i]<-toclean$meanin[i]
  }
}
activityclean<-toclean
rm(toclean)
```

## Total steps every day for imputed dataset
```{r total steps imputed}
steps2<-with(activityclean,tapply(steps,date,sum, na.rm=T))
hist(steps2, xlab= "Total Steps- imputed", main= " Total Steps Taken Every Day-imputed", col= "steel blue")
mean2<-round(mean(steps))
median2<-median(steps)
changemean<-((mean2-mean0)/mean0)*100
changemedian<-((median2-median0)/median0)*100
```
The median of total steps taken in a day is `r median2`, mean is `r mean2`.
%change in median is `r changemedian` and %change in mean is `r changemean`.
## variation between weekdays and weekends
```{r}
activityclean$day<-strftime(activityclean$date, "%A")
i<-0
while(i<nrow(activityclean)){
  i<-i+1
  if(activityclean$day[i] =="Monday"){
    activityclean$type[i]<-"Weekday"
  }
  if(activityclean$day[i] =="Tuesday"){
    activityclean$type[i]<-"Weekday"
  }
  if(activityclean$day[i] =="Wednesday"){
    activityclean$type[i]<-"Weekday"
  }
  if(activityclean$day[i] =="Thursday"){
    activityclean$type[i] <-"Weekday"
  }
  if(activityclean$day[i] =="Friday"){
    activityclean$type[i]<-"Weekday"
  }
  else{
    activityclean$type[i] <- "Weekend"
  }
}
weekday<-subset(activityclean,activityclean$type == "Weekday")
weekend<-subset(activityclean,activityclean$type == "Weekend")
weekdayaverage<-aggregate(steps~interval, weekday, FUN = mean)
weekendaverage<-aggregate(steps~interval, weekend, FUN = mean)
par(mfrow=c(2,1))
with(weekdayaverage,plot(interval,steps,xlab = "Interval",type="l",col="orange", main =   "Average for weekdays", ylab = "Average Steps"))
with(weekendaverage,plot(interval,steps,xlab = "Interval",type="l",col="orange",main = "Average for weekends", ylab= "Average Steps"))
```
